How Tree-like are the data?

### 2019-05-02 phyloglmm_ms.tex
Do we care about the magic theshold?

Unlike standard statistical models, where all of the predictor variables of interest (for example, species traits and environmental factors) are directly observable,  PCAs use phylogenetic relationships to estimate the amount of unobserved process of trait evolution (i.e how "tree-like" are the observed data given the phylogenetic tree) \citep{felsenstein1985phylogenies, butler2004phylogenetic}. 

@@

@@

>>>>

Unlike standard statistical models, where all of the predictor variables of interest (for example, species traits and environmental factors) are directly observable,  PCAs use phylogenetic relationships to estimate the amount of unobserved process of trait evolution (i.e how ``tree-like" are the observed data given the phylogenetic tree) \citep{felsenstein1985phylogenies, butler2004phylogenetic}. 

@@

\subsection*{How ``tree-like" are the data?}

Multispecies data sets are often very challenging to model using basic statistical frameworks due to the complexity and non-normality data. 
Most researchers typically use basic regression based methods such as generalized-linear models (GLMs) models and suffocated ecologist would use generalized-linear mixed effect models \citep{bolker2009generalized} to control for different levels of variation.
Most of the basic statistical techniques assume species are independent, and indentically distributed (if species are not accounted for in the model), which are false from a evolutionary perspective.
However, traits

@@

\mli{this part doesn't make sense. Going back to the tree-like example, we can generate two test: cor vs no cor and tip vs noise}

### 2019-05-02 phylotreelike.rmd
Do we care about the magic theshold?

```{r, sampling, echo=FALSE,cache=TRUE}
ddprop <- data.frame(nspp=3:100, prop=NA)
rep_pp <- pp
for(i in 3:100){
phy <- rtree(n = i)
Vphy <- vcv(phy)

@@

@@

dd$null <- sapply(1:nrow(dd),function(x){mvtnorm::dmvnorm(dd[x,1:i], mean =rep(0,i), sigma = Vnull, log = TRUE)})

dd2 <- (dd
  %>% rowwise()
  %>% mutate(LR = ll-null
             # , LRT = max(as.numeric(LR > log(c)),0.5)
             , LRT = abs(LR) > c/2
             )
  # %>% arrange(LR)
)

ddprop$prop[i-2] <- mean(dd2$LRT)
}

plot(ddprop)

@@

ggprop <- (ggplot(ddprop, aes(x=nspp,y=prop))
  + geom_point()
  + geom_line()
  + theme_bw()
  + ylab("Proportion of Tree-like traits")
  + xlab("Number of species")
)

>>>>

```{r, sampling, echo=FALSE,cache=TRUE}
ddprop <- data.frame(nspp=3:50, cortest=NA, nulltest=NA, startest=NA)
rep_pp <- pp
for(i in 3:50){
phy <- rtree(n = i)
Vphy <- vcv(phy)

wrong_tree <- rtree(n = i)
wrongVphy <- vcv(wrong_tree)

@@

dd$wll <- sapply(1:nrow(dd),function(x){mvtnorm::dmvnorm(dd[x,1:i], mean =rep(0,i), sigma = wrongVphy, log = TRUE)})

@@

dd$null <- sapply(1:nrow(dd),function(x){mvtnorm::dmvnorm(dd[x,1:i], mean =rep(0,i), sigma = Vnull, log = TRUE)})
dd$null2 <- sapply(1:nrow(dd),function(x){logLik(lm(unlist(dd[x,1:i])~-1))})

dd2 <- (dd
  %>% rowwise()
  %>% mutate(LR = ll-null
             # , LRT = max(as.numeric(LR > log(c)),0.5)
             , LR2 = ll-null2
             , LR3 = null-null2
             , LRT = ll > null # abs(LR) > c/2
             , LRT1 = ll > wll
             , LRT2 = ll > null2 # abs(LR2) > (qchisq(0.95,df=(2*i-1))/2)
             , LRT3 = null > null2 # abs(LR3) > (qchisq(0.95,df=(i-1))/2)
             , LRT4 = wll > null
             , LRT5 = wll > null2
             )
  # %>% arrange(LR)
)

ddprop$tree_cor[i-2] <- mean(dd2$LRT)
ddprop$wrong_tree[i-2] <- mean(dd2$LRT1)
ddprop$tree_test[i-2] <- mean(dd2$LRT2)
ddprop$star_test[i-2] <- mean(dd2$LRT3)
ddprop$wrongtree_cor[i-2] <- mean(dd2$LRT4)
ddprop$wrongtree_test[i-2] <- mean(dd2$LRT5)
}

@@

ddprop2 <- (ddprop
  %>% gather(key="LRT", value="proportion", -nspp)
  %>% na.omit()
)

ggprop <- (ggplot(ddprop2, aes(x=nspp,y=proportion,color=LRT))
  + geom_point()
  + geom_line()
  + theme_bw()
  + ylab("Proportion of Tree-like traits")
  + xlab("Number of species")
)
